---
title: "OTZ 3PM Report"
author: "Dominic Ngalo"
date: "February 13, 2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

This is an R Script to populate the data.Before you Run the script, *please indicate the period of interest for reporting here below:*

```{r}
fromDate <- as.Date("2019-10-01")
toDate <- as.Date("2019-10-31")
```



### Importing the data in .csv format

```{r data importation}
library(knitr)
library(ggplot2)
library(gridExtra)
library('RODBC')
library(dplyr)
library(ggthemes)


```

## Top ten list

We can view the top 10 list of clients currently in otz!

```{r pressure, echo=FALSE}
otzRegister <- read.csv("otzregister.csv", header = TRUE, sep = ",")
dim(otzRegister)

```


## Baseline Information

### Total ALHIV enrolled in OTZ during the month
```{r Enrolled, echo=FALSE}

otzRegister$OTZStartDate <- as.Date(otzRegister$OTZStartDate, origin = "1899-12-30")
otzRegister$BaselineVLDate <- as.Date(otzRegister$BaselineVLDate, origin = "1899-12-30")
newEnrolled <- subset(otzRegister, otzRegister$OTZStartDate > fromDate & otzRegister$OTZStartDate < toDate)
with(newEnrolled, table(Gender))

```

### ALHIV in OTZ with baseline VL results (VL within the last 6 months) at enrolment into OTZ

```{r Baseline Viral Load, echo=FALSE}
newWithVL <- subset(newEnrolled,  !is.null(newEnrolled$BaselineVL))
with(newWithVL, table(Gender))
```

### ALHIV 's enrolled in OTZ with VL< 1000 at enrolment into OTZ

```{r with suppressed Baseline VL <1000 , echo=FALSE}
suppressedBaseline <- subset(newWithVL, newWithVL$BaselineVL < 1000)
with(suppressedBaseline, table(Gender))

```

### ALHIV 's enrolled in OTZ with VL<400 at enrolment into OTZ

```{r , echo=FALSE}
suppressedBaseline <- subset(newWithVL, newWithVL$BaselineVL < 400)
with(suppressedBaseline, table(Gender))
```

### ALHIV 's enrolled in OTZ with VL= LDL at enrolment into OTZ

```{r , echo=FALSE}
suppressedBaseline <- subset(newWithVL, newWithVL$BaselineVL < 1)
with(suppressedBaseline, table(Gender))
```

## Additional monitoring

### Number of adolescents in OTZ who were booked for appointments in the month

```{r booked for Appointment, echo=FALSE}
otzRegister$NextAppointmentDate <- as.Date(otzRegister$NextAppointmentDate, origin = "1899-12-30")
otzRegister$LastVisitDate <- as.Date(otzRegister$LastVisitDate, origin = "1899-12-30")


booked_1 <- subset(otzRegister, otzRegister$NextAppointmentDate >= fromDate & otzRegister$NextAppointmentDate <=toDate )
booked_2 <- subset(otzRegister, otzRegister$LastVisitDate >= fromDate & otzRegister$LastVisitDate <=toDate )


appointed <- rbind(booked_1, booked_2)

with(appointed, table(Gender))

```

### Number of adolescents in OTZ who kept (came on/before the scheduled date) their clinic appointments
```{r}
with(booked_2, table(Gender))
```

### Number of adolescents in OTZ with adherence > 95% adherence

```{r}
adherent <- subset(booked_2, booked_2$LatestVL < 1000)
with(adherent, table(Gender))
```

### Number of OTZ members who attended support group
```{r}
with(booked_2, table(Gender))
```


# COHORT REPORTING [10 - 19 years]

```{r}
'%notin%' <- Negate('%in%')
cohort_1a <- subset(otzRegister, otzRegister$AgeCurrent >= 10 & otzRegister$AgeCurrent <= 19)
cohort_1 <- subset(cohort_1a, cohort_1a$ARTOutcome %notin% c('Transfer', 'Death', 'Lost', 'Refused Care', 'Transfer'))
cohort_1$LatestVLDate <- as.Date(cohort_1$LatestVLDate, origin = "1899-12-30")
cohort_1$WithRoutine <- ifelse((cohort_1$LatestVLDate >= (fromDate-60) & cohort_1$LatestVLDate <=toDate), "Yes", "No")
cohort_1$Unsuppressed <- ifelse((cohort_1$LatestVL > 1000 & cohort_1$WithRoutine == "Yes"), "Yes", "No")
cohort_1$Suppressed <- ifelse((cohort_1$LatestVL < 1000 & cohort_1$WithRoutine == "Yes"), "Yes", "No")
cohort_1$Less400 <- ifelse((cohort_1$LatestVL < 400 & cohort_1$WithRoutine == "Yes"), "Yes", "No")
cohort_1$LDL <- ifelse((cohort_1$LatestVL < 1 & cohort_1$WithRoutine == "Yes"), "Yes", "No")
```

## Routine VL monitoring to assess the VL Testing Uptake and suppression

### Total ALHIV in OTZ who were eligible for routine viral load testing during the reporting period

#### MONTH 6

```{r}
eligible6 <- subset(cohort_1, cohort_1$OTZStartDate >= (fromDate - 180) & cohort_1$OTZStartDate >= (toDate - 180))
with(eligible6, table(Gender))
```

#### MONTH 12

```{r}
eligible12 <- subset(cohort_1, cohort_1$OTZStartDate >= (fromDate - 360) & cohort_1$OTZStartDate >= (toDate - 360))
with(eligible12, table(Gender))
```

#### MONTH 18

```{r}
eligible18 <- subset(cohort_1, cohort_1$OTZStartDate >= (fromDate - 540) & cohort_1$OTZStartDate >= (toDate - 540))
with(eligible18, table(Gender))

```

#### MONTH 24

```{r}
eligible24 <- subset(cohort_1, cohort_1$OTZStartDate >= (fromDate - 720) & cohort_1$OTZStartDate >= (toDate - 720))
with(eligible24, table(Gender))
```

### Number of ALHIV in OTZ whose samples were taken for routine viral load testing
```{r}
takenEligible6 <- subset(eligible6, eligible6$WithRoutine == 'Yes')
takenEligible12 <- subset(eligible12, eligible12$WithRoutine == 'Yes')
takenEligible18 <- subset(eligible18, eligible18$WithRoutine == 'Yes')
takenEligible24 <- subset(eligible24, eligible24$WithRoutine == 'Yes')

```


#### MONTH 6

```{r}
with(takenEligible6, table(Gender))
```


#### MONTH 12
```{r}
with(takenEligible12, table(Gender))
```


#### MONTH 18
```{r}
with(takenEligible18, table(Gender))
```


#### MONTH 24
```{r}
with(takenEligible24, table(Gender))
```



### Number of ALHIV in OTZ with routine follow up VL results at the end of the reporting month
#### MONTH 6

```{r}
with(takenEligible6, table(Gender))
```


#### MONTH 12
```{r}
with(takenEligible12, table(Gender))
```


#### MONTH 18
```{r}
with(takenEligible18, table(Gender))
```


#### MONTH 24
```{r}
with(takenEligible24, table(Gender))
```

### Number with follow up VL > 1000 copies/ml
```{r}
Over1000Eligible6 <- subset(eligible6, eligible6$Unsuppressed == 'Yes')
Over1000Eligible12 <- subset(eligible12, eligible12$Unsuppressed == 'Yes')
Over1000Eligible18 <- subset(eligible18, eligible18$Unsuppressed == 'Yes')
Over1000Eligible24 <- subset(eligible24, eligible24$Unsuppressed == 'Yes')
```

#### MONTH 6
```{r}
with(Over1000Eligible6, table(Gender))
```


#### MONTH 12
```{r}
with(Over1000Eligible12, table(Gender))
```

#### MONTH 18
```{r}
with(Over1000Eligible18, table(Gender))
```


#### MONTH 24

```{r}
with(Over1000Eligible24, table(Gender))
```

### Number with follow up VL < 1000 copies/ml
### Number of ALHIV in OTZ whose samples were taken for routine viral load testing
```{r}
less1000Eligible6 <- subset(eligible6, eligible6$Suppressed == 'Yes')
less1000Eligible12 <- subset(eligible12, eligible12$Suppressed == 'Yes')
less1000Eligible18 <- subset(eligible18, eligible18$Suppressed == 'Yes')
less1000Eligible24 <- subset(eligible24, eligible24$Suppressed == 'Yes')

```

#### MONTH 6
```{r}
with(less1000Eligible6, table(Gender))
```


#### MONTH 12
```{r}
with(less1000Eligible12, table(Gender))
```

#### MONTH 18
```{r}
with(less1000Eligible18, table(Gender))
```


#### MONTH 24

```{r}
with(less1000Eligible24, table(Gender))
```


### Number with routine follow up VL < 400 copies/ml
```{r}
less400Eligible6 <- subset(eligible6, eligible6$Less400 == 'Yes')
less400Eligible12 <- subset(eligible12, eligible12$Less400 == 'Yes')
less400Eligible18 <- subset(eligible18, eligible18$Less400 == 'Yes')
less400Eligible24 <- subset(eligible24, eligible24$Less400 == 'Yes')


```

#### MONTH 6
```{r}
with(less400Eligible6, table(Gender))
```


#### MONTH 12
```{r}
with(less400Eligible12, table(Gender))
```


#### MONTH 18
```{r}
with(less400Eligible18, table(Gender))
```

#### MONTH 24
```{r}
with(less400Eligible24, table(Gender))
```

### Number with routine follow up VL = LDL
```{r}
LDLEligible6 <- subset(eligible6, eligible6$LDL == 'Yes')
LDLEligible12 <- subset(eligible12, eligible12$LDL == 'Yes')
LDLEligible18 <- subset(eligible18, eligible18$LDL == 'Yes')
LDLEligible24 <- subset(eligible24, eligible24$LDL == 'Yes')
```


#### MONTH 6
```{r}
with(LDLEligible6, table(Gender))
```

#### MONTH 12
```{r}
with(LDLEligible12, table(Gender))
```

#### MONTH 18
```{r}
with(LDLEligible18, table(Gender))
```

#### MONTH 24
```{r}
with(LDLEligible24, table(Gender))
```

##Routine VL monitoring to assess actual overall suppression for the cohort

### Total ALHIV in OTZ who had valid viral load results during the reporting period
```{r}
overallCohort <- subset(cohort_1, !is.null(cohort_1$LatestVL))
#with(overallCohort, table(Gender))
```


#### MONTH 6
```{r}
overall6 <- subset(overallCohort, (overallCohort$OTZStartDate >= (toDate - 180) & overallCohort$OTZStartDate <=(toDate)))
with(overall6, table(Gender))
```


#### MONTH 12
```{r}
overall12 <- subset(overallCohort, (overallCohort$OTZStartDate >= (toDate - 360) & overallCohort$OTZStartDate <=(toDate -180)))
with(overall12, table(Gender))
```


#### MONTH 18
```{r}
overall18 <- subset(overallCohort, (overallCohort$OTZStartDate >= (toDate - 540) & overallCohort$OTZStartDate <=(toDate -360)))
with(overall18, table(Gender))
```


#### MONTH 24

```{r}
overall24 <- subset(overallCohort, (overallCohort$OTZStartDate >= (toDate - 720) & overallCohort$OTZStartDate <=(toDate -540)))
with(overall24, table(Gender))
```


### Total ALHIV in OTZ who had valid viral load results > 1000 copies/ml

#### MONTH 6
```{r}
with((subset(overall6, overall6$LatestVL > '1000')),table(Gender))
```

#### MONTH 12
```{r}
with((subset(overall12, overall12$LatestVL > '1000')),table(Gender))
```

#### MONTH 18
```{r}
with((subset(overall18, overall18$LatestVL > '1000')),table(Gender))
```

#### MONTH 24
```{r}
with((subset(overall24, overall24$LatestVL > '1000')),table(Gender))
```


### Total ALHIV in OTZ who had valid viral load results < 1000 copies/ml

#### MONTH 6
```{r}
with((subset(overall6, overall6$LatestVL < 1000)),table(Gender))
```

#### MONTH 12
```{r}
with((subset(overall12, overall12$LatestVL < 1000)),table(Gender))
```

#### MONTH 18
```{r}
with((subset(overall18, overall18$LatestVL < 1000)),table(Gender))
```

#### MONTH 24
```{r}
with((subset(overall24, overall24$LatestVL < 1000)),table(Gender))
```


### Total ALHIV in OTZ who had valid viral load results < 400 copies/ml

#### MONTH 6
```{r}
with((subset(overall6, overall6$LatestVL < 400)),table(Gender))
```

#### MONTH 12
```{r}
with((subset(overall12, overall12$LatestVL < 400)),table(Gender))
```

#### MONTH 18
```{r}
with((subset(overall18, overall18$LatestVL < 400)),table(Gender))
```

#### MONTH 24
```{r}
with((subset(overall24, overall24$LatestVL < 400)),table(Gender))
```


### Total ALHIV in OTZ who had valid viral load results = LDL

#### MONTH 6
```{r}
with((subset(overall6, overall6$LatestVL < 1)),table(Gender))
```

#### MONTH 12
```{r}
with((subset(overall12, overall12$LatestVL < 1)),table(Gender))
```

#### MONTH 18
```{r}
with((subset(overall18, overall18$LatestVL < 1)),table(Gender))
```

#### MONTH 24
```{r}
with((subset(overall24, overall24$LatestVL < 1)),table(Gender))
```

## Tracking those with suspected treatment failure at baseline or within the reporting period

### Total ALHIV in OTZ who had valid viral load results > 1000 copies/ml
```{r}
highBaseline <- subset(cohort_1, cohort_1$BaselineVL > 1000)
```
#### MONTH 6
```{r}
highB6 <- subset(highBaseline, ((OTZStartDate >= (toDate-180)) & (OTZStartDate <= (toDate-0))))
with(highB6, table(Gender))
```

#### MONTH 12
```{r}
highB12 <- subset(highBaseline, ((OTZStartDate >= (toDate-360)) & (OTZStartDate <= (toDate-180))))
with(highB12, table(Gender))
```
#### MONTH 18
```{r}
highB18 <- subset(highBaseline, ((OTZStartDate >= (toDate-540)) & (OTZStartDate <= (toDate-360))))
with(highB18, table(Gender))
```
#### MONTH 24
```{r}
highB24 <- subset(highBaseline, ((OTZStartDate >= (toDate-720)) & (OTZStartDate <= (toDate-540))))
with(highB24, table(Gender))
```

### Total ALHIV for the period of review who had repeat VL test results

#### MONTH 6
```{r}
with(highB6, table(Gender))
```

#### MONTH 12
```{r}
with(highB12, table(Gender))
```
#### MONTH 18
```{r}
with(highB18, table(Gender))
```
#### MONTH 24
```{r}
with(highB24, table(Gender))
```

### Number with repeat VL < 1000 copies/ml

#### MONTH 6
```{r}
with(subset(highB6, highB6$LatestVL < 1000), table(Gender))
```

#### MONTH 12
```{r}
with(subset(highB12, highB12$LatestVL < 1000), table(Gender))
```

#### MONTH 18
```{r}
with(subset(highB18, highB18$LatestVL < 1000), table(Gender))
```

#### MONTH 24
```{r}
with(subset(highB24, highB24$LatestVL < 1000), table(Gender))
```


### Number with repeat VL < 400 copies/ml
#### MONTH 6
```{r}
with(subset(highB6, highB6$LatestVL < 400), table(Gender))
```
#### MONTH 12
```{r}
with(subset(highB12, highB12$LatestVL < 400), table(Gender))
```
#### MONTH 18
```{r}
with(subset(highB18, highB18$LatestVL < 400), table(Gender))
```
#### MONTH 24
```{r}
with(subset(highB24, highB24$LatestVL < 400), table(Gender))
```

### Number with repeat VL = LDL
#### MONTH 6
```{r}
with(subset(highB6, highB6$LatestVL < 1), table(Gender))
```
#### MONTH 12
```{r}
with(subset(highB12, highB12$LatestVL < 1), table(Gender))
```
#### MONTH 18
```{r}
with(subset(highB18, highB18$LatestVL < 1), table(Gender))
```
#### MONTH 24
```{r}
with(subset(highB24, highB24$LatestVL < 1), table(Gender))
```
### Number with repeat VL > 1000 copies/ml
#### MONTH 6
```{r}
with(subset(highB6, highB6$LatestVL > 1000), table(Gender))
```

#### MONTH 12
```{r}
with(subset(highB12, highB12$LatestVL > 1000), table(Gender))
```

#### MONTH 18
```{r}
with(subset(highB18, highB18$LatestVL > 1000), table(Gender))
```

#### MONTH 24
```{r}
with(subset(highB24, highB24$LatestVL > 1000), table(Gender))
```

### Number switched to second line ART
#### MONTH 6
```{r}
with(subset(highB6, highB6$CurrentRegimenLine == 2), table(Gender))
```

#### MONTH 12
```{r}
with(subset(highB12, highB12$CurrentRegimenLine == 2), table(Gender))
```
#### MONTH 18
```{r}
with(subset(highB18, highB18$CurrentRegimenLine == 2), table(Gender))
```
#### MONTH 24
```{r}
with(subset(highB24, highB24$CurrentRegimenLine == 2), table(Gender))
```

### Number switched to third line ART
#### MONTH 6
```{r}
with(subset(highB6, highB6$CurrentRegimenLine == 3), table(Gender))
```

#### MONTH 12
```{r}
with(subset(highB12, highB12$CurrentRegimenLine == 3), table(Gender))
```
#### MONTH 18
```{r}
with(subset(highB18, highB18$CurrentRegimenLine == 3), table(Gender))
```
#### MONTH 24
```{r}
with(subset(highB24, highB24$CurrentRegimenLine == 3), table(Gender))
```

## Tracking attritions

```{r}

cohort_Attrition <- subset(otzRegister, otzRegister$AgeCurrent >= '10' & otzRegister$AgeCurrent <= '19')


attrition6 <- subset(cohort_Attrition, (OTZStartDate >= (toDate - 180) & OTZStartDate <=(toDate)))
attrition12 <- subset(cohort_Attrition, (OTZStartDate >= (toDate - 360) & OTZStartDate <=(toDate-180)))
attrition18 <- subset(cohort_Attrition, (OTZStartDate >= (toDate - 540) & OTZStartDate <=(toDate-360)))
attrition24 <- subset(cohort_Attrition, (OTZStartDate >= (toDate - 720) & OTZStartDate <=(toDate-540)))

```


### Number transferred out during the review period
#### MONTH 6
```{r}
with(subset(attrition6, ARTOutcome =='Transfer'), table(Gender))
```

#### MONTH 12
```{r}
with(subset(attrition12, ARTOutcome =='Transfer'), table(Gender))
```

#### MONTH 18
```{r}
with(subset(attrition18, ARTOutcome =='Transfer'), table(Gender))
```

#### MONTH 24
```{r}
with(subset(attrition24, ARTOutcome =='Transfer'), table(Gender))

```

### Number lost to follow up this during the review period
#### MONTH 6
```{r}
with(subset(attrition6, ARTOutcome =='Lost'), table(Gender))
```

#### MONTH 12
```{r}
with(subset(attrition12, ARTOutcome =='Lost'), table(Gender))
```

#### MONTH 18
```{r}
with(subset(attrition18, ARTOutcome =='Lost'), table(Gender))
```

#### MONTH 24
```{r}
with(subset(attrition24, ARTOutcome =='Lost'), table(Gender))

```
### Number aged 20-24 years and transitioned to adult care during the review period 
#### MONTH 6
#### MONTH 12
#### MONTH 18
#### MONTH 24
### Number reported as dead during the review period
#### MONTH 6
```{r}
with(subset(attrition6, ARTOutcome =='Death'), table(Gender))
```

#### MONTH 12
```{r}
with(subset(attrition12, ARTOutcome =='Death'), table(Gender))
```

#### MONTH 18
```{r}
with(subset(attrition18, ARTOutcome =='Death'), table(Gender))
```

#### MONTH 24
```{r}
with(subset(attrition24, ARTOutcome =='Death'), table(Gender))

```
### Number that opted out of OTZ during the review period
#### MONTH 6
```{r}
with(subset(attrition6, ARTOutcome =='Refused Care'), table(Gender))
```

#### MONTH 12
```{r}
with(subset(attrition12, ARTOutcome =='Refused Care'), table(Gender))
```

#### MONTH 18
```{r}
with(subset(attrition18, ARTOutcome =='Refused Care'), table(Gender))
```

#### MONTH 24
```{r}
with(subset(attrition24, ARTOutcome =='Refused Care'), table(Gender))

```


# COHORT REPORTING [20 - 24 years]
```{r}
'%notin%' <- Negate('%in%')
cohort_2a <- subset(otzRegister, otzRegister$AgeCurrent >= 20 & otzRegister$AgeCurrent <= 24)
cohort_2 <- subset(cohort_2a, ARTOutcome %notin% c('Transfer', 'Death', 'Lost', 'Refused Care', 'Transfer'))


yp6 <- subset(cohort_2, (OTZStartDate >= (toDate-180) & OTZStartDate <=(toDate)))
yp12 <- subset(cohort_2,(OTZStartDate >=(toDate-360) & OTZStartDate <=(toDate-180)))
yp18 <- subset(cohort_2, (OTZStartDate >=(toDate-540) & OTZStartDate <=(toDate-360)))
yp24 <- subset(cohort_2, (OTZStartDate >=(toDate-720) & OTZStartDate <=(toDate-540)))

```



## Continuing services

### Number of youths and young persons aged 20-24 years and still in the program
#### MONTH 6
```{r}
with(yp6, table(Gender))
```

#### MONTH 12
```{r}
with(yp12, table(Gender))
```

#### MONTH 18
```{r}
with(yp18, table(Gender))
```

#### MONTH 24
```{r}
with(yp24, table(Gender))
```
### Total youths aged 20-24 years and still in OTZ who had valid viral load results during the reporting period
#### MONTH 6
```{r}
with(yp6, table(Gender))
```

#### MONTH 12
```{r}
with(yp12, table(Gender))
```

#### MONTH 18
```{r}
with(yp18, table(Gender))
```

#### MONTH 24
```{r}
with(yp24, table(Gender))
```
### Total youths aged 20-24 years in OTZ who had valid viral load results < 1000 copies/ml
#### MONTH 6
```{r}
with(subset(yp6, LatestVL < 1000),table(Gender))
```

#### MONTH 12
```{r}
with(subset(yp12, LatestVL < 1000),table(Gender))
```
#### MONTH 18
```{r}
with(subset(yp18, LatestVL < 1000),table(Gender))
```
#### MONTH 24
```{r}
with(subset(yp24, LatestVL < 1000), table(Gender))
```
### Total youths aged 20-24 years in OTZ who had valid viral load results < 400 copies/ml
#### MONTH 6
```{r}
with(subset(yp6, LatestVL < 400),table(Gender))
```

#### MONTH 12
```{r}
with(subset(yp12, LatestVL < 400),table(Gender))
```
#### MONTH 18
```{r}
with(subset(yp18, LatestVL < 400),table(Gender))
```
#### MONTH 24
```{r}
with(subset(yp24, LatestVL < 400), table(Gender))
```
### Number of youths aged 20-24 years who exited from Post OTZ group during the review period
#### MONTH 6
#### MONTH 12
#### MONTH 18
#### MONTH 24
